---
title: "Course 6, project part 2"
output: html_document
author: Feng Li
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is the second portion of the course 6 project, we're going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data and perform some basic exploratory data analysis
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

## Load data and perform basic exploratory data analysis

```{r}
library(datasets)
data(ToothGrowth)

dim(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)

summary(ToothGrowth)
```
And we'll do some plots for a better view.

```{r}
library(ggplot2)
g <- ggplot(ToothGrowth, aes(supp, len)) +
    geom_boxplot(aes(fill=dose)) + xlab("Dose Amount") + ylab("Tooth Length") + facet_grid(~ supp) +     ggtitle("Tooth Length vs. Dose Amount \nby Delivery Method") + 
    theme(plot.title = element_text(lineheight=.8, face="bold"))

g
```







